---
title: "Arithmetic Complexity Filter of 6/xx Lotto tickets"
author: "John Akwei, ECMp ERMp, Founder of ContextBase"
date: "Saturday, August 01, 2015"
output: html_document
---
  
## Synopsis:  
Arithmetic complexity shows the variations of differences between the numbers of analyzed tickets.  The higher the complexity, the more likely the number is characteristic of a random number generated by the Lotto number generator. A human creating a Lotto ticket number usually creates a number with a lower complexity than a random number generator.  

#### The formula is:  
Arithmetic Complexity(t) = D(t) - (r-1) where D(t) is the count of unique differences between all ticket numbers, and r is the count of numbers in the analyzed ticket.  

In a 6/xx lottery the Arithmetic Complexity value ranges from 0 to 10. About 74% of all possible combinations have the Arithmetic Complexity value in the range of 8 to 10.  

#### How ticket number differences are calculated:  
Let's analyze ticket 1, 2, 3, 4, 5, 6. Differences are calculated via:  
2-1 = 1  
3-1 = 2  
4-1 = 3  
5-1 = 4  
6-1 = 5  
3-2 = 1  
4-2 = 2  
5-2 = 3  
6-2 = 4  
4-3 = 1  
5-3 = 2  
6-3 = 3  
5-4 = 1  
6-4 = 2  
6-5 = 1  

There are 5 unique differences: 1, 2, 3, 4, 5 so the Arithmetic Complexity of this ticket is 5-(6-1) = 0.  

#### To demonstrate the algorithm, the following datasets are used:   
```{r}
load("intermediate.rdata")
Table20 <- read.csv("Table_20_rows.csv")
TableToAdd <- read.csv("Table_to_add.csv")
```

### The first 6 Lotto ticket numbers:  
```{r}
head(Table20)
```

#### A dataset of 1,473,817 Lotto ticket numbers:  
```{r}
str(data)
```

#### A function I authored to find the differences in 6/xx Lotto numbers:  
```{r}
ticketDiff <- function(ticket) {
x <- 1
y <- 1
z <- 5
for(i in 1:5) {
for(i in 1:z) {
  diff <- ticket[1, x + y] - ticket[1, x]
  print(diff)
  y <- y + 1 }
x <- x + 1
y <- 1
z <- z - 1 } }
```

### Applying the ticketDiff function to the first ticket:  
```{r}
# The last winning Lotto ticket
head(Table20, 1)
# Differences in the numbers of the Lotto ticket
ticketDiff(Table20)
```

### Applying the ticketDiff function to the second ticket:  
```{r}
# The second winning Lotto ticket in the data table
head(Table20[2,], 1)
# Differences in the above ticket number
ticketDiff(Table20[2, ])
```

### Code to fill in new 6/xx Lotto numbers:  
```{r}
NewLottoNum <- matrix(nrow=1, ncol=6)
NewLottoNum[] <- c(2, 4, 23, 23, 38, 46)
NewLottoNum <- data.frame(NewLottoNum)
names(NewLottoNum) <- c("V1", "V2", "V3", "V4", "V5", "V6")
ticketDiff(NewLottoNum)
```

### Calculating the Arithmetic Complexity Number:  
```{r}
uq <- unique(c(10, 17, 18, 20, 24, 7, 8, 10, 14, 1, 3, 7, 2, 6, 4))
Dt <- length(uq)
acNum <- Dt - (6 - 1)
```

### The amount of unique difference numbers:  
```{r}
Dt
```

### The Arithmetic Complexity Number of the Lotto ticket:  
```{r}
acNum
```

#### Converting intermediate.rdata  
```{r, eval=F}
newData <- data.frame(data)
interData <- matrix(nrow=1473817, ncol=6)
interData <- data.frame(interData)
names(interData) <- c("V1","V2","V3","V4","V5","V6")
y <- 1
for(i in 1:1473817) {
  interData[y,] <- newData[,y]
  y <- y + 1 }
```

```{r, echo=F}
newData <- data.frame(data)
interData <- matrix(nrow=100, ncol=6)
interData <- data.frame(interData)
names(interData) <- c("V1","V2","V3","V4","V5","V6")
y <- 1
for(i in 1:100) {
  interData[y,] <- newData[,y]
  y <- y + 1 }
head(interData)
```

#### Finding Table20 Matching 6/xx numbers in intermediate.rdata  
```{r, eval=F}
m <- Table20[1:20, ]
res2 <- rowSums(matrix(m %in% interData, ncol=ncol(m)))
```

```{r, echo=F}
m <- Table20[1:20, ]
res2 <- rowSums(matrix(m %in% interData, ncol=ncol(m)))
```

#### Matches found in datatable of lotto numbers:
```{r}
res2
```

## Summary  
Unfortunately, the more valuable a Lottery prize is, the less likely the Lottery is honest. There are extensive connections between Mafia organized gangs, Governments, Police organizations, the Pentagon, etc. The priorities of Lottery organizations might include Security, Patriotism, economic control, etc.  

From the Washington Post, http://www.washingtonpost.com/blogs/ezra-klein/post/how-to-hack-the-lottery/2011/08/24/gIQA3yZHbJ_blog.html:  

"Most lotteries, obviously, want to turn a profit, €”especially if they'€™re supposed to provide a relatively stable source of revenue for the government. And that means they want to control how many people win".  

Also, "The game can'€™t be truly random. Instead, it has to generate the illusion of randomness while actually being carefully determined".€

Data Science Central examined predicting winning lottery numbers in 2011 - http://www.analyticbridge.com/profiles/blogs/are-lottery-winning-numbers-really-random.  

Also, here is some programming help - http://www.analyticbridge.com/profiles/blogs/how-to-detect-a-pattern-problem-and-solution.  